\bf Baseline Drift Estimation for Time Series Data Using Quantile Trend Filtering

### 2019-02-08 Manuscript/detrendify_JCGS.tex
More of Eric's edits.

	\begin{align*}
	y_1(t) = y(t) & \mbox{~~if~~} 1 \le t \le u_{1}\\
	y_2(t) = y(t) & \mbox{~~if~~} l_{2} \le t \le u_{2} \\
	\vdots & \\
	y_M(t) = y(t) & \mbox{~~if~~} l_{M} \le t \le  n,
	\end{align*}
with boundaries $1 < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < \cdots< n$. An illustration is given in \Fig{windows}. Given quantiles $\tau_1 < \ldots < \tau_J$ to be estimated we define $\theta_{j,m}(t)$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $M$ at time point $t$. In order to write out the constraint that the overlapping sections must be equal, we define a consensus variable
	 \begin{align*}
		 \bar{\theta}_{j,m} =&  g(\theta_{j, m-1}, \theta_{j,m}, \theta_{j,m+1}) \\
		 =& \begin{cases} 
			 \frac{\theta_{j,m-1}(t)+\theta_{j,m}(t)}{2} & \mbox{if~~} l_{m} \le t \le u_{m-1}  \\
			 \theta_{j,m}(t) & \mbox{if~~} u_{m-1} \le t \le l_{m+1}  \\
			 \frac{\theta_{j,m}(t)+\theta_{j,m+1}(t)}{2} & \mbox{if~~} l_{m+1} \le t \le u_{m}  
			 \end{cases},
	\end{align*}
	defining $\theta_{j,M+1} = \theta_{j,M}$ and $\theta_{j,0} = \theta_{j,1}$. Our windowed quantile trend optimization problem can then be written as 
	 \begin{align*}
		 \label{eq:quantile_windows}
		 &\underset{\theta_{1,m}, \ldots, \theta_{J,m}}{\min}\; \sum_{m=1}^M\sum_{j=1}^J \left [\rho_{\tau_j}(y_m - \theta_{j,m}) + 
		 \lambda_j \lVert \Mn{D}{k+1} \theta_{j,m} \rVert_1 \right ] \\
		 &\text{subject to: }\; \\
		 &\qquad \theta_{1,m}(t) \le \theta_{2,m}(t) \le \ldots \le \theta_{J,m}(t) \text{ for all } m,t \\
		 &\qquad \theta_{j,m}(t) = \bar{\theta}_{j,m}(t) \text{ for all } j, m, t
	 \end{align*}
	 Defining Lagrange multipliers $\omega_{j,m}$, and penalty parameter $\gamma > 0$ we can write the augmented Lagrangian for finding the trends in window $m$:
	 \begin{equation*}
	 \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert \Mn{D}{k+1}\theta_{j,m}\rVert_1 +  \omega_{j,m}\Tra(\theta_{j,m} - \bar{\theta}_{j,m}) + 
	 \frac{\gamma}{2}||\theta_{j,m} - \bar{\theta}_{j,m}||_2^2
	 \end{equation*}
	 We then estimate the trend separately in each window, which can be done in parallel, while constraining the overlapping pieces of the trends to be equal as outlined in algorithm 1.

	\begin{figure}[!h] 
		\centering
		\caption{Window boundaries and trends fit separately in each window.}
		\includegraphics[width = 0.8\linewidth]{Figures/overlapping_windows.png}
		\label{fig:windows}
	\end{figure}

@@

	\subsection{Regularization Parameter Choice}
	\label{sec:lambda_choice}
	An important problem in trend estimation is the choice of regularization parameter or degree of smoothness. Our method can easily handle missing data by defining the check-loss function to output 0 for missing values. This allows us to leave out validation observations that can be used to select the tuning parameter $\lambda$. However, the use of an information criteria metric can result in a better choice of regularization parameter than the validation method.  \cite{KoenkerNgPortnoy1994} addressed the choice of regularization parameter by proposing the Schwarz criterion for the selection of $\lambda$
	\begin{eqnarray*}
	\label{eq:SIC}
	\mbox{SIC}(p_{\lambda}) & = & \log\left[\frac{1}{n}\rho_{\tau}(y - \theta)\right] + \frac{1}{2n}p_{\lambda}\log n.
	\end{eqnarray*}
	where $p_{\lambda} = \sum_t \One(y(t) = \widehat{\theta}(t))$ is the number of interpolated points, which can be thought of as active knots. The SIC is based on the traditional Bayesian Information Criterion (BIC) which is given by 
	\begin{eqnarray*}
	\mbox{BIC}(s) & = & -2\log(L\{\hat{\theta}\}) + \nu\log n 
	\end{eqnarray*}	
	where $L$ is the likelihood function and $\nu$ is the number of non-zero components in $\hat{\theta}$. If we take the approach used in Bayesian quantile regression \citep{yu2001bayesian}, and assume that minimizing the check-loss function corresponds to maximizing the asymmetric Laplace likelihood, 
	\begin{eqnarray*}
	L(y \mid \theta) & = & \left(\frac{\tau^n(1-\tau)}{\sigma}\right)^n\exp\left\{-\sum_t\rho_\tau\left(\frac{y(t) - \theta(t)}{\sigma}\right)\right\},
	\end{eqnarray*} 
	we can compute the BIC as 
	\begin{eqnarray*}
	\mbox{BIC}(\mbox{df}) & = & 2\frac{1}{\sigma}\rho_{\tau}(y-\hat{\theta}) + \mbox{df}\log n
	\end{eqnarray*} 
	where $df$ is the number of non-zero elements of $\Mn{D}{k+1}\Vhat{\theta}$. We can choose any $\sigma>0$ and have found empirically that $\sigma =  \frac{1-\lvert 1-2\tau\rvert}{2}$ produces stable estimates.

>>>>

	\begin{align*}
	y_1(t) = y(t) & \mbox{~~if~~} 1 \le t \le u_{1}\\
	y_2(t) = y(t) & \mbox{~~if~~} l_{2} \le t \le u_{2} \\
	\vdots & \\
	y_M(t) = y(t) & \mbox{~~if~~} l_{M} \le t \le  n,
	\end{align*}
with boundaries $1 < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < \cdots< n$. \Fig{windows} illustrates how we break the samples into overlapping windows. Given quantiles $\tau_1 < \ldots < \tau_J$ to be estimated we define $\theta_{j,m}(t)$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $M$ at time point $t$. In order to write out the constraint that the overlapping sections must be equal, we define a consensus variable
	 \begin{align*}
		 \bar{\theta}_{j,m} =&  g(\theta_{j, m-1}, \theta_{j,m}, \theta_{j,m+1}) \\
		 =& \begin{cases} 
			 \frac{\theta_{j,m-1}(t)+\theta_{j,m}(t)}{2} & \mbox{if~~} l_{m} \le t \le u_{m-1}  \\
			 \theta_{j,m}(t) & \mbox{if~~} u_{m-1} \le t \le l_{m+1}  \\
			 \frac{\theta_{j,m}(t)+\theta_{j,m+1}(t)}{2} & \mbox{if~~} l_{m+1} \le t \le u_{m}  
			 \end{cases},
	\end{align*}
	defining $\theta_{j,M+1} = \theta_{j,M}$ and $\theta_{j,0} = \theta_{j,1}$. Our windowed quantile trend optimization problem can then be written as 
	 \begin{align*}
		 \label{eq:quantile_windows}
		 &\underset{\theta_{1,m}, \ldots, \theta_{J,m}}{\text{minimize}}\; \sum_{m=1}^M\sum_{j=1}^J \left [\rho_{\tau_j}(y_m - \theta_{j,m}) + 
		 \lambda_j \lVert \Mn{D}{k+1} \theta_{j,m} \rVert_1 \right ] \\
		 &\text{subject to: }\; \\
		 &\qquad \theta_{1,m}(t) \le \theta_{2,m}(t) \le \ldots \le \theta_{J,m}(t) \text{ for all } m,t \\
		 &\qquad \theta_{j,m}(t) = \bar{\theta}_{j,m}(t) \text{ for all } j, m, t
	 \end{align*}
	 Defining Lagrange multipliers $\omega_{j,m}$, and penalty parameter $\gamma > 0$ we can write the augmented Lagrangian for finding the trends in window $m$:
	 \begin{equation*}
	 \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert \Mn{D}{k+1}\theta_{j,m}\rVert_1 +  \omega_{j,m}\Tra(\theta_{j,m} - \bar{\theta}_{j,m}) + 
	 \frac{\gamma}{2}||\theta_{j,m} - \bar{\theta}_{j,m}||_2^2
	 \end{equation*}
	 We then estimate the trend separately in each window, which can be done in parallel, while constraining the overlapping pieces of the trends to be equal as outlined in algorithm 1.

	\begin{figure}[!h] 
		\centering
		\includegraphics[width = 0.8\linewidth]{Figures/overlapping_windows.png}
		\caption{Window boundaries and trends fit separately in each window.}		
		\label{fig:windows}
	\end{figure}

@@

	\subsection{Regularization Parameter Choice}
	\label{sec:lambda_choice}
	An important problem in trend estimation is the choice of regularization parameter or degree of smoothness. Our method can easily handle missing data by defining the check-loss function to output 0 for missing values. This allows us to leave out validation observations that can be used to select the tuning parameter $\lambda$. However, the use of an information criteria metric can result in a better choice of regularization parameter than the validation method.  \cite{KoenkerNgPortnoy1994} addressed the choice of regularization parameter by proposing the Schwarz criterion for the selection of $\lambda$
	\begin{eqnarray*}
	\label{eq:SIC}
	\mbox{SIC}(p_{\lambda}) & = & \log\left[\frac{1}{n}\rho_{\tau}(y - \theta)\right] + \frac{1}{2n}p_{\lambda}\log n.
	\end{eqnarray*}
	where $p_{\lambda} = \sum_t \One(y(t) = \widehat{\theta}(t))$ is the number of interpolated points, which can be thought of as active knots. The SIC is based on the traditional Bayesian Information Criterion (BIC) which is given by 
	\begin{eqnarray*}
	\mbox{BIC}(s) & = & -2\log(L\{\hat{\theta}\}) + \nu\log n 
	\end{eqnarray*}	
	where $L$ is the likelihood function and $\nu$ is the number of non-zero components in $\hat{\theta}$. \Halley{Eric}{What is $s$?} If we take the approach used in Bayesian quantile regression \citep{yu2001bayesian}, and assume that minimizing the check-loss function corresponds to maximizing the asymmetric Laplace likelihood, 
	\begin{eqnarray*}
	L(y \mid \theta) & = & \left(\frac{\tau^n(1-\tau)}{\sigma}\right)^n\exp\left\{-\sum_t\rho_\tau\left(\frac{y(t) - \theta(t)}{\sigma}\right)\right\},
	\end{eqnarray*} 
	we can compute the BIC as 
	\begin{eqnarray*}
	\mbox{BIC}(\mbox{df}) & = & 2\frac{1}{\sigma}\rho_{\tau}(y-\hat{\theta}) + \mbox{df}\log n
	\end{eqnarray*} 
	where $df$ is the number of non-zero elements of $\Mn{D}{k+1}\Vhat{\theta}$.  \Halley{Eric}{What is $\hat{\theta}$? This is the first time you used this notation.} We can choose any $\sigma>0$ and have found empirically that $\sigma =  \frac{1-\lvert 1-2\tau\rvert}{2}$ produces stable estimates.
